---
title: "gbif_occs"
output: html_document
date: "2026-02-03"
---

# Pull GBIF / iDigBio occurrence data

To get started, we're going to be pulling code from the [Soltis lab
ENM Botany 2025 workshop](https://soltislab.github.io/BotanyENMWorkshops/Botany2025/Botany2025.html). 

## package management

### download packages

```{r eval = FALSE}
library(devtools)
list_of_packages <- c(
  # Core ENM tools
  "terra", "dismo", "ENMeval", "ENMTools", "biomod2",
  # Spatial / Mapping
  "sf", "rnaturalearth", "ggspatial", "leaflet",
  "fields", "rangeBuilder",
  # Visualization
  "ggplot2", "viridis", "gridExtra",
  # Data manipulation
  "dplyr", "tidyr", "stringr", "gtools",
  # Phylogenetics
  "ape", "phytools",
  # Biodiversity data
  "ridigbio", "gatoRs",
  # Other utilities
  "multcompView", "usdm", "predicts", "rJava"
)

new.packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]

if (length(new.packages)) {
  cat("\nInstalling missing CRAN packages:\n")
  print(new.packages)
  install.packages(new.packages)
} else {
  cat("\nAll required CRAN packages already installed.\n")
}

## ---- Load All CRAN Packages ----

# Try loading all required CRAN packages.
loaded <- sapply(list_of_packages, require, character.only = TRUE)

if (any(!loaded)) {
  cat("\nWARNING: These CRAN packages failed to load:\n")
  print(list_of_packages[!loaded])
} else {
  cat("\nAll CRAN packages loaded successfully.\n")
}

github_repos <- c(
  "vqv/ggbiplot"
)

for (repo in github_repos) {
  pkg <- tail(strsplit(repo, "/")[[1]], 1)
  if (!pkg %in% installed.packages()[,"Package"]) {
    cat(paste("\nInstalling GitHub package:", repo, "\n"))
    devtools::install_github(repo, upgrade = "never")
  } else {
    cat(paste("\nGitHub package already installed:", pkg, "\n"))
  }
}

# Try loading GitHub packages.
github_packages <- c("ggbiplot")
github_loaded <- sapply(github_packages, require, character.only = TRUE)

if (any(!github_loaded)) {
  cat("\nWARNING: These GitHub packages failed to load:\n")
  print(github_packages[!github_loaded])
} else {
  cat("\nAll GitHub packages loaded successfully.\n")
}
```

### load packages - R
```{r}
library(ridigbio)       # Interface to iDigBio API
library(gatoRs)         # Taxonomic and geographic cleaning
library(fields)         # For calculating spatial distances
library(sf)             # For spatial operations and visualization
library(ggplot2)        # For plotting
library(ggspatial)      # Scale bars and north arrows
```

### load packages - julia
```{julia}
using DataFrames, RCall
using CSV
using JLD2
using Plots
using GeoIO
# Load required R packages
R"""
library(gatoRs)
library(ggplot2)
library(sf)
library(ggspatial)
library(gridExtra)
library(CoordinateCleaner)
"""
```

## practice pulling occurrences for just 2 taxa

```{r}
#I don't think I need to add the synonyms that are already listed under GBIF, maybe just the synonyms
    #that aren't -- those that are in Kew/POWO?
Calceolaria_corymbosa <- c("Calceolaria corymbosa",
                            "Calceolaria glabrata var. meyeniana")
Calceolaria_thyrsiflora <- c("Calceolaria thyrsiflora", 
                         "Fagelia thyrsiflora",
                         "Calceolaria thyrsiflora var. dulcis",
                         "Calceolaria thyrsiflora var. alliacea")

gators_download(synonyms.list = Calceolaria_corymbosa,
                write.file = TRUE,
                filename = "data/occurrence_data/pt_occs_raw/corymbosa_2026_02_03.csv")

gators_download(synonyms.list = Calceolaria_thyrsiflora,
                write.file = TRUE,
                filename = "data/occurrence_data/pt_occs_raw/thyrsiflora_2026_02_03.csv")

rawdf1 <- read.csv("data/occurrence_data/pt_occs_raw/corymbosa_2026_02_03.csv")
rawdf2 <- read.csv("data/occurrence_data/pt_occs_raw/thyrsiflora_2026_02_03.csv")
names(rawdf1) # returns field headers         
nrow(rawdf1) # returns the number of rows = occurrence records   

leaflet(rawdf2) %>% 
  addTiles() %>% 
  addMarkers(label = paste0(rawdf$longitude,
                            ", ",
                            rawdf$latitude))
```

```{r}
unique(rawdf1$scientificName)

# Clean names with fuzzy matching
df <- taxa_clean(df = rawdf1,
                 synonyms.list = Calceolaria_corymbosa,
                 taxa.filter = "fuzzy", # fuzzy allows some leeway over exact matching
                 accepted.name = "Calceolaria corymbosa")
#1286 points before
#if we select exact, WAAAAAAY fewer points , like ~80

df <- basic_locality_clean(df = df,
                           remove.zero = TRUE,
                           precision = TRUE,
                           digits = 2,
                           remove.skewed = TRUE)

#884 points after

df <- process_flagged(df, 
                      interactive = FALSE, 
                      scientific.name = "accepted_name")

nrow(df) #786 points

df <- remove_duplicates(df, 
                        remove.unparseable = TRUE)

nrow(df) #783 points

df <- one_point_per_pixel(df)
nrow(df) #570
```

### practice plotting

```{r}
nnDm <- rdist.earth(as.matrix(df[, c("longitude", "latitude")]),
                    miles = FALSE)
nnDmin <- do.call(rbind, 
                  lapply(1:5, function(i) sort(nnDm[, i])[2]))
min(nnDmin)  # e.g., 0.933 km

df_fixed <- st_as_sf(df, 
                     coords = c("longitude", "latitude"), 
                     crs = 4326)

world <- annotation_borders(database = "world", 
               colour = "gray80", 
               fill = "gray80")

simple_map <- ggplot() +
                world + 
                geom_sf(data = df_fixed, 
                        color = "blue") +
                coord_sf(xlim = c(min(df$longitude) - 3, 
                                  max(df$longitude) + 3),
                         ylim = c(min(df$latitude) - 3,
                                  max(df$latitude) + 3)) +
                xlab("Longitude") + 
                ylab("Latitude") +
                annotation_scale() +
                annotation_north_arrow(location = "tl", 
                                       height = unit(1, "cm"),
                                       width = unit(1, "cm"))

write.csv(df, "data/occurrence_data/pt_occs_clean/corymbosa_2026_02_03_cleaned.csv", row.names = FALSE)

leaflet(df_fixed) %>%
  addMarkers(label = paste0(df$longitude, ", ", df$latitude)) %>%
  addTiles()
```

## pull occurrences for all taxa

```{julia}
# load in dictionary
synonymdict = load("data/taxonomy_trait_data/synonymdict.jld2")
# load in dataframe
taxa_traits = DataFrame(CSV.File("data/taxonomy_trait_data/traits_species_cleaned-2026_02_25.csv"))

# Track failures and no-data taxa
failed_taxa = String[]
successful_taxa = String[]
no_data_taxa = String[]

for taxa in keys(synonymdict)
    synlist = [syn[1] for syn in synonymdict[taxa]]
    filename = replace(taxa, " " => "_")
    filepath = "data/occurrence_data/pt_occs_raw/$(filename)-2026_02_18.csv"
    
    # Check if file already exists
    if isfile(filepath)
        println("  Skipping - file already exists")
        push!(successful_taxa, taxa)
        continue
    end
    
    @rput synlist
    @rput filepath
    
    try
        R"gators_download(synonyms.list = synlist, write.file = TRUE, filename = filepath)"
        println("  Success!")
        push!(successful_taxa, taxa)
        sleep(3)
        
    catch e
        error_msg = string(e)
        
        if occursin("No records found", error_msg)
            println("  No occurrence data found for this taxon")
            push!(no_data_taxa, taxa)
            sleep(1)
        else
            println("  ERROR: $e")
            push!(failed_taxa, taxa)
            sleep(10)
        end
    end
end

if !isempty(no_data_taxa)
    println("\nTaxa with no occurrence records:")
    for taxa in no_data_taxa
        println("  - $taxa")
    end
end

if !isempty(failed_taxa)
    println("\nFailed taxa (errors):")
    for taxa in failed_taxa
        println("  - $taxa")
    end
end
```

no failed taxa, no no-data taxa. cool!

## clean occurrence data

batch cleaning for all species:
```{julia}
# Load homonyms file to check for bad authorities
homonyms = DataFrame(CSV.File("data/taxonomy_trait_data/homonyms.csv"))
false_entries = homonyms[homonyms.accepted .== false, :]
true_entries = homonyms[homonyms.accepted .== true, :]

# Create a dictionary for quick lookup: taxon name -> bad authority
bad_authorities = Dict{String, String}()
for row in eachrow(false_entries)
    bad_authorities[row.scientificName] = row.authority
end

# Create reverse lookup: 
# If validName from FALSE entry is in our dataset, we need to filter OUT the TRUE entry
# BUT only if validName != scientificName (skip self-referential cases)
# Example: "Calceolaria glandulosa Benth." (FALSE) points to "Calceolaria nitida"
#          "Calceolaria nitida" IS in our dataset
#          So when processing "Calceolaria nitida", filter OUT "Calceolaria glandulosa Poepp. ex Benth." (a valid name!)
authorities_to_exclude = Dict{String, Vector{String}}()  # validName -> [authorities to exclude]

dataset_taxa = Set(keys(synonymdict))

for row in eachrow(false_entries)
    valid_name = row.validName
    sci_name = row.scientificName
    
    # Skip if no validName
    if ismissing(valid_name) || valid_name == ""
        continue
    end

    # Skip self-referential cases (validName == scientificName)
    if valid_name == sci_name
        continue
    end
    
    # Check if this validName is in our dataset
    if valid_name in dataset_taxa
        # Find the TRUE entry for this scientificName
        true_match = filter(r -> r.scientificName == sci_name && r.accepted == true, homonyms)
        
        if nrow(true_match) > 0
            true_authority = true_match[1, :authority]
            
            # When processing validName, exclude the TRUE authority
            if !haskey(authorities_to_exclude, valid_name)
                authorities_to_exclude[valid_name] = String[]
            end
            push!(authorities_to_exclude[valid_name], true_authority)
        end
    end
end

println("Taxa with bad authorities to filter: $(length(bad_authorities))")
println("Taxa needing reverse exclusion: $(length(authorities_to_exclude))")
for (valid_name, exclude_auths) in authorities_to_exclude
    for auth in exclude_auths
        println("  $valid_name will EXCLUDE authority: '$auth'")
    end
end
```

```{julia}
# Initialize tracking dataframe
filtering_stats = DataFrame(
    taxa = String[],
    raw = Int[],
    after_taxa_clean = Int[],
    after_remove_duplicates = Int[],
    removed_bad_authorities = Int[],
    removed_reverse_exclusions = Int[],
    rows_missing_coords = Int[],
    rows_with_coords = Int[],
    after_locality_filt = Int[],
    after_remove_flagged = Int[]
    #after_1ptper_pixel = Int[]
)

for taxa in keys(synonymdict)    
    # get the synonym list for this species
    synlist = [syn[1] for syn in synonymdict[taxa]]
    filename = replace(taxa, " " => "_")
    filepath = "data/occurrence_data/pt_occs_raw/$(filename)-2026_02_18.csv"
    
    # Check if this taxon has a bad authority to filter
    bad_auth = get(bad_authorities, taxa, nothing)
    
    # Check if this taxon needs reverse exclusion (filter out TRUE authorities)
    exclude_auths = get(authorities_to_exclude, taxa, String[])

    # pass to R
    @rput filepath
    @rput synlist
    @rput bad_auth
    @rput exclude_auths
    
    # run R cleaning pipeline with row tracking
    R"""
    df <- read.csv(filepath)
    nrow_raw <- nrow(df)
    
    df <- taxa_clean(df,
                     synonyms.list = synlist, 
                     taxa.filter = "fuzzy",
                     accepted.name = synlist[1])
    nrow_taxa_clean <- nrow(df)
    
    # remove duplicates between GBIF and iDigBio
    df <- remove_duplicates(df, 
                            remove.unparseable = TRUE)
    nrow_remove_dup <- nrow(df)

    # Filter out rows with bad authorities (for homonym cases)
    nrow_removed_bad_auth <- 0
    if (!is.null(bad_auth) && bad_auth != "") {
        df_before <- nrow(df)
        df <- df[!grepl(bad_auth, df$scientificName, fixed = TRUE), ]
        nrow_removed_bad_auth <- df_before - nrow(df)
    }
    
    # Filter out rows with reverse exclusion authorities
    # Example: for Calceolaria nitida, exclude "Poepp. ex Benth." 
    nrow_removed_reverse <- 0
    if (length(exclude_auths) > 0) {
        df_before <- nrow(df)
        for (auth in exclude_auths) {
            df <- df[!grepl(auth, df$scientificName, fixed = TRUE), ]
        }
        nrow_removed_reverse <- df_before - nrow(df)
    }
    
    # Separate rows with NA lat/long for potential georeferencing
    df_na <- df[is.na(df$latitude) | is.na(df$longitude), ]
    # filter out rows where absolutely no locality info
    df_na <- df_na[df_na$locality != "locality:  NA, occurrenceRemarks: NA, verbatimLocality: NA", ]
    df_clean <- df[!is.na(df$latitude) & !is.na(df$longitude), ]
    
    nrow_na <- nrow(df_na)
    nrow_with_coords <- nrow(df_clean)
    
    df_clean <- basic_locality_clean(df_clean, 
                                remove.zero = TRUE, 
                                precision = TRUE, 
                                digits = 2, 
                                remove.skewed = TRUE)
    nrow_locality <- nrow(df_clean)
    
    # remove points above or below lat/long bounding box
    df_clean <- df_clean %>%
         filter(latitude  >= -56,  latitude  <= 31.5,
                longitude >= -117, longitude <= -33)
    # some parts of texas fall into this range above so let's manually remove USA pts
    df_clean <- df_clean %>% filter(country != "United States of America")
    #fixit: incorporate native range polygons?
    #df_clean <- clean_coordinates(df_clean,
    #                              lon = "longitude", lat = "latitude", 
    #                              species = "scientificName",
    #                              value = "clean")
    nrow_flagged <- nrow(df_clean)
    
    #df_clean <- one_point_per_pixel(df_clean)
    #nrow_pixel <- nrow(df_clean)
    """
    
    # Get row counts from R
    @rget nrow_raw
    @rget nrow_taxa_clean
    @rget nrow_remove_dup
    @rget nrow_removed_bad_auth
    @rget nrow_removed_reverse
    @rget nrow_na
    @rget nrow_with_coords
    @rget nrow_locality
    @rget nrow_flagged
    #@rget nrow_pixel
    
    # add to tracking dataframe
    push!(filtering_stats, (taxa, nrow_raw, nrow_taxa_clean, nrow_remove_dup, 
                            nrow_removed_bad_auth, nrow_removed_reverse,
                            nrow_na, nrow_with_coords, nrow_locality, nrow_flagged))
    
    println("  $taxa: Raw=$nrow_raw → Taxa clean=$nrow_taxa_clean → Remove dup=$nrow_remove_dup")
    if nrow_removed_bad_auth > 0
        println("    ⚠️  Removed $nrow_removed_bad_auth rows with bad authority: '$bad_auth'")
    end
    if nrow_removed_reverse > 0
        println("    ⚠️  Removed $nrow_removed_reverse rows with excluded authorities: $(join(exclude_auths, ", "))")
    end
    println("    Split: $nrow_with_coords with coords, $nrow_na missing coords")
    
    # save cleaned df with coordinates
    outfile = "data/occurrence_data/pt_occs_clean/$(filename)-2026_02_18_cleaned.csv"
    @rput outfile
    R"write.csv(df_clean, outfile, row.names = FALSE)"
    
    # save rows with NA coordinates (if any exist)
    if nrow_na > 0
        outfile_na = "data/occurrence_data/pt_occs_to_georeference/$(filename)-2026_02_18_no_coords.csv"
        @rput outfile_na
        R"write.csv(df_na, outfile_na, row.names = FALSE)"
    end
    
    R"rm(df, df_clean, df_na, outfile)"
end
```

### check minimum number of occurrences
```{julia}
# add calculated columns for seeing filtering before and after nrow
filtering_stats.taxa_clean_removed = filtering_stats.raw .- filtering_stats.after_taxa_clean
filtering_stats.duplicates_removed = filtering_stats.after_taxa_clean .- filtering_stats.after_remove_duplicates
filtering_stats.total_removed = filtering_stats.raw .- filtering_stats.after_remove_duplicates
filtering_stats.percent_retained = round.((filtering_stats.after_remove_duplicates ./ filtering_stats.raw) .* 100, digits=1)

# show taxa with filtering
filtered_taxa = filtering_stats[filtering_stats.total_removed .> 0, :]
println("Taxa with records removed ($(nrow(filtered_taxa)) taxa):")
println(filtered_taxa)

# save the stats
#CSV.write("data/filtering_stats_2026_02_18.csv", filtering_stats)
#println("\nFiltering statistics saved to data/filtering_stats_2026_02_18.csv")
```

let's look through some of these:
1. Calceolaria alba
> this is because many rows were Calceolaria thyrsiflora for some reason? can't see why. so good filtering here.
   
2. Calceolaria collina subsp. conferta
> removing Calceolaria fulva f. viscosa --> good filtering, this is a synonym of Calceolaria fulva.

3. Calceolaria corymbosa subsp. corymbosa
> removes TONS of Calceolaria corymbosa (without any infraspecific label), so good job.

4. Calceolaria uniflora
> removes many Calceolaria virgata (which is a valid taxon), so good job.

let's see how many taxa do not have enough occurrences.

```{julia}
toosmall = nrowcomp[ ( nrowcomp.nrow_before .< 25), :]
histogram(nrowcomp.nrow_before)
```

create priority list for georeferencing:
```{julia}
# sort descending on rows missing coords (prioritizing rows w/ many we can georeference)
sort!(filtering_stats, :rows_missing_coords, rev=true)
# sort ascending on rows with very few coords as is (w/ lat/longs)
sort!(filtering_stats, :after_locality_filt, rev=false)
# separate out those where the sum of :rows_missing_coords + :after_locality_filt < 15 (not as much of a priority)
toofewobs = filter(row -> row.rows_missing_coords + row.after_locality_filt < 15, filtering_stats)
enoughobs = filter(row -> row.rows_missing_coords + row.after_locality_filt >= 15, filtering_stats)
CSV.write("data/occurrence_data/ranked_togeoref.csv", enoughobs)
```

## visualize occurrences before & after filtering

```{julia}
# Get list of taxa from synonymdict
taxa_list = taxa_traits.scientificName

# Load TDWG botanical country polygons once
R"bot_regions = st_read('data/occurrence_data/bot_country_shapefiles/level3.shp', quiet = TRUE)"

# Build lookup of taxon -> botanical country codes from taxa_traits
taxa_to_nativerange_dict = Dict()
for row in eachrow(taxa_traits)
    taxa_to_nativerange_dict[row.scientificName] = row.botanicalCountries
end

plotted_count = 0

# Create PDF
pdf_file = "data/occurrence_data/occurrence_maps_before_after.pdf"
R"pdf($pdf_file, width = 14, height = 7)"

for (idx, taxa) in enumerate(taxa_list)
    println("Processing $idx/$(length(taxa_list)): $taxa")
    
    filename = replace(taxa, " " => "_")
    raw_file = "data/occurrence_data/pt_occs_raw/$(filename)-2026_02_18.csv"
    clean_file = "data/occurrence_data/pt_occs_clean/$(filename)-2026_02_18_cleaned.csv"
    
    # Check if files exist
    if !isfile(raw_file) || !isfile(clean_file)
        println("  Skipping - files not found")
        continue
    end
    
    # Read the data
    raw_df = DataFrame(CSV.File(raw_file))
    clean_df = DataFrame(CSV.File(clean_file))
    
    # Filter out NA coordinates for raw data
    raw_df = filter(row -> !ismissing(row.latitude) && !ismissing(row.longitude), raw_df)
    
    # Skip if no data
    if nrow(raw_df) == 0 || nrow(clean_df) == 0
        println("  Skipping - no valid coordinates")
        continue
    end
    
    # Pass data to R
    @rput taxa
    @rput raw_file
    @rput clean_file
    country_codes = get(taxa_to_nativerange_dict, taxa, String[])
    @rput country_codes
    
    R"""
    # Read data
    raw_df <- read.csv(raw_file)
    clean_df <- read.csv(clean_file)
    
    # Remove NA coordinates
    raw_df <- raw_df[!is.na(raw_df$latitude) & !is.na(raw_df$longitude), ]
    
    # Skip if empty
    if (nrow(raw_df) == 0 || nrow(clean_df) == 0) {
        next
    }
    
    # Calculate bounds for raw data (includes all points)
    raw_lon_min <- min(raw_df$longitude, na.rm = TRUE) - 3
    raw_lon_max <- max(raw_df$longitude, na.rm = TRUE) + 3
    raw_lat_min <- min(raw_df$latitude, na.rm = TRUE) - 3
    raw_lat_max <- max(raw_df$latitude, na.rm = TRUE) + 3
    
    # Calculate bounds for cleaned data (tighter bounds based on cleaned points only)
    clean_lon_min <- min(clean_df$longitude, na.rm = TRUE) - 2
    clean_lon_max <- max(clean_df$longitude, na.rm = TRUE) + 2
    clean_lat_min <- min(clean_df$latitude, na.rm = TRUE) - 2
    clean_lat_max <- max(clean_df$latitude, na.rm = TRUE) + 2
    
    # Get world and country borders
    world <- annotation_borders(database = "world", 
                                 colour = "gray80", 
                                 fill = "gray80")
    
    countries <- annotation_borders(database = "world",
                                    colour = "gray40",
                                    fill = NA,
                                    size = 0.5)
    
    # Taxon-specific botanical country polygons (if available)
    bot_underlay <- bot_regions[bot_regions$LEVEL3_COD %in% country_codes, ]
    underlay_layer <- if (nrow(bot_underlay) > 0) {
        geom_sf(data = bot_underlay,
                fill = "khaki1",
                color = "goldenrod4",
                alpha = 0.25,
                linewidth = 0.3)
    } else {
        NULL
    }

    # Create before map
    p1 <- ggplot() +
        world +
        countries +
        underlay_layer +
        geom_point(data = raw_df, 
                   aes(x = longitude, y = latitude),
                   color = "blue",
                   size = 1.5,
                   alpha = 0.6) +
        coord_sf(xlim = c(raw_lon_min, raw_lon_max),
                 ylim = c(raw_lat_min, raw_lat_max)) +
        labs(title = paste0("Raw (n=", nrow(raw_df), ")"),
             x = "Longitude",
             y = "Latitude") +
        theme_minimal() +
        theme(plot.title = element_text(size = 10))
    
    # Create after map
    p2 <- ggplot() +
        world +
        countries +
        underlay_layer +
        geom_point(data = clean_df, 
                   aes(x = longitude, y = latitude),
                   color = "darkgreen",
                   size = 1.5,
                   alpha = 0.6) +
        coord_sf(xlim = c(clean_lon_min, clean_lon_max),
                 ylim = c(clean_lat_min, clean_lat_max)) +
        labs(title = paste0("Cleaned (n=", nrow(clean_df), ")"),
             x = "Longitude",
             y = "Latitude") +
        annotation_scale(location = "bl") +
        annotation_north_arrow(location = "tl", 
                               height = unit(0.8, "cm"),
                               width = unit(0.8, "cm")) +
        theme_minimal() +
        theme(plot.title = element_text(size = 10))
    
    # Combine plots
    combined <- grid.arrange(p1, p2, ncol = 2,
                             top = taxa)
    
    print(combined)
    """
    
    plotted_count += 1
end

# Close PDF
R"dev.off()"
```

```{julia}
function filter_countries(taxon::String, countries::Vector{String})
    # fixit: could make this smarter and not hard code the date at the end and instead do some kind of string matching thing
    path = "data/occurrence_data/pt_occs_clean/" * replace(taxon, " " => "_") * "-2026_02_18_cleaned.csv"
    cleandf = DataFrame(CSV.File(path))
    # check that countries are even in df
    dfcountries = cleandf.country
    if length(intersect(countries, dfcountries)) == 0
        error("no input country provided in `countries` list matches countries existing in dataframe")
    end

    # filter every country out of df
    for country in countries
        filter!(:country => x -> x != country, cleandf)
    end
    # write out CSV again
    CSV.write(path, cleandf)
end
```


problems observed in maps:
Ehrhart 2005 (Sys. Bot.):
1. integrifolia - this manuscript defines / carves out several species from integrifolia s.l.,
    defining integrifolia s.s. might be a problem if all these occurrences haven't been annotated / revisited post 2005.
    but Peru can definitely be removed.

```{julia}
filter_countries("Calceolaria integrifolia", ["Peru"])
```

Ehrhart monograph:
2. ascendens subsp. glandulifera
fixit: deal with nivalis (get rid of all, but essentially deal with synonym problem)
3. ascendens subsp. ascendens
4. polifolia
5. nitida
6. cana - need to remove the Mexican point
```{julia}
filter_countries("Calceolaria cana", ["Mexico"])
```

7. purpurea
8. corymbosa subsp. corymbosa
9.  corymbosa subsp. montana 
10. glandulosa subsp. glandulosa
11. crenatiflora
12. filicaulis subsp. filicaulis
13. uniflora
14. biflora
15. undulata
16. tenella
17. hypericina
18. dentata subsp. dentata

Molau:
19. inflexa - Molau says endemic to Peru. I don't know if I believe the two Bolivian points.
20. umbellata - fixit: delete the -35ºS pt, it comes from GenBank.
21. speciosa - fixit: need to dig into this data, US herbarium had their specimens digitized wrong. 
        should be endemic to Cuzco departimiento, per Molau.
22. angustiflora - matching Molau, if not missing some in southern Peru.
23. chrysophaera - two points in Bolivia, which doesn't agree with Molau. but those two were det. by Puppo, who I trust.

## fixits

- edge cases for members in my `homonyms.csv` file --> 
```{julia}
homonymsa = intersect(taxa_traits.scientificName, homonyms.scientificName)
homonymsb = intersect(taxa_traits.scientificName, homonyms.validName)
homonyms_indataset = union(homonymsa, homonymsb)
```

there's 13 of these, we'll need to be extra careful and double check:
  1. occurrences of the synonym/homonym without species authority
  2. look at overlap of range between synonyms

- cleaning coords:
  - remove human obs? do something with trusted iNat observers?
  - filter using native range botanical country polygons +/- some buffer?
- get summary stats for cultivated vs. inat vs. etc
- after georeferencing, figure out reporting back to herbaria?
- update priority list for Nikitha