---
title: "gbif_occs"
output: html_document
date: "2026-02-03"
---

# Pull GBIF / iDigBio occurrence data

To get started, we're going to be pulling code from the [Soltis lab
ENM Botany 2025 workshop](https://soltislab.github.io/BotanyENMWorkshops/Botany2025/Botany2025.html). 

## package management

### download packages

```{r eval = FALSE}
library(devtools)
list_of_packages <- c(
  # Core ENM tools
  "terra", "dismo", "ENMeval", "ENMTools", "biomod2",
  # Spatial / Mapping
  "sf", "rnaturalearth", "ggspatial", "leaflet",
  "fields", "rangeBuilder",
  # Visualization
  "ggplot2", "viridis", "gridExtra",
  # Data manipulation
  "dplyr", "tidyr", "stringr", "gtools",
  # Phylogenetics
  "ape", "phytools",
  # Biodiversity data
  "ridigbio", "gatoRs",
  # Other utilities
  "multcompView", "usdm", "predicts", "rJava"
)

new.packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]

if (length(new.packages)) {
  cat("\nInstalling missing CRAN packages:\n")
  print(new.packages)
  install.packages(new.packages)
} else {
  cat("\nAll required CRAN packages already installed.\n")
}

## ---- Load All CRAN Packages ----

# Try loading all required CRAN packages.
loaded <- sapply(list_of_packages, require, character.only = TRUE)

if (any(!loaded)) {
  cat("\nWARNING: These CRAN packages failed to load:\n")
  print(list_of_packages[!loaded])
} else {
  cat("\nAll CRAN packages loaded successfully.\n")
}

github_repos <- c(
  "vqv/ggbiplot"
)

for (repo in github_repos) {
  pkg <- tail(strsplit(repo, "/")[[1]], 1)
  if (!pkg %in% installed.packages()[,"Package"]) {
    cat(paste("\nInstalling GitHub package:", repo, "\n"))
    devtools::install_github(repo, upgrade = "never")
  } else {
    cat(paste("\nGitHub package already installed:", pkg, "\n"))
  }
}

# Try loading GitHub packages.
github_packages <- c("ggbiplot")
github_loaded <- sapply(github_packages, require, character.only = TRUE)

if (any(!github_loaded)) {
  cat("\nWARNING: These GitHub packages failed to load:\n")
  print(github_packages[!github_loaded])
} else {
  cat("\nAll GitHub packages loaded successfully.\n")
}
```

### load packages - R
```{r}
library(ridigbio)       # Interface to iDigBio API
library(gatoRs)         # Taxonomic and geographic cleaning
library(fields)         # For calculating spatial distances
library(sf)             # For spatial operations and visualization
library(ggplot2)        # For plotting
library(ggspatial)      # Scale bars and north arrows
```

### load packages - julia
```{julia}
using DataFrames, RCall
using CSV
using JLD2
using Plots
R"library(gatoRs)"
```

## practice pulling occurrences for just 2 taxa

```{r}
#I don't think I need to add the synonyms that are already listed under GBIF, maybe just the synonyms
    #that aren't -- those that are in Kew/POWO?
Calceolaria_corymbosa <- c("Calceolaria corymbosa",
                            "Calceolaria glabrata var. meyeniana")
Calceolaria_thyrsiflora <- c("Calceolaria thyrsiflora", 
                         "Fagelia thyrsiflora",
                         "Calceolaria thyrsiflora var. dulcis",
                         "Calceolaria thyrsiflora var. alliacea")

gators_download(synonyms.list = Calceolaria_corymbosa,
                write.file = TRUE,
                filename = "data/pt_occs_raw/corymbosa_2026_02_03.csv")

gators_download(synonyms.list = Calceolaria_thyrsiflora,
                write.file = TRUE,
                filename = "data/pt_occs_raw/thyrsiflora_2026_02_03.csv")

rawdf1 <- read.csv("data/pt_occs_raw/corymbosa_2026_02_03.csv")
rawdf2 <- read.csv("data/pt_occs_raw/thyrsiflora_2026_02_03.csv")
names(rawdf1) # returns field headers         
nrow(rawdf1) # returns the number of rows = occurrence records   

leaflet(rawdf2) %>% 
  addTiles() %>% 
  addMarkers(label = paste0(rawdf$longitude,
                            ", ",
                            rawdf$latitude))
```

```{r}
unique(rawdf1$scientificName)

# Clean names with fuzzy matching
df <- taxa_clean(df = rawdf1,
                 synonyms.list = Calceolaria_corymbosa,
                 taxa.filter = "fuzzy", # fuzzy allows some leeway over exact matching
                 accepted.name = "Calceolaria corymbosa")
#1286 points before
#if we select exact, WAAAAAAY fewer points , like ~80

df <- basic_locality_clean(df = df,
                           remove.zero = TRUE,
                           precision = TRUE,
                           digits = 2,
                           remove.skewed = TRUE)

#884 points after

df <- process_flagged(df, 
                      interactive = FALSE, 
                      scientific.name = "accepted_name")

nrow(df) #786 points

df <- remove_duplicates(df, 
                        remove.unparseable = TRUE)

nrow(df) #783 points

df <- one_point_per_pixel(df)
nrow(df) #570
```

### practice plotting

```{r}
nnDm <- rdist.earth(as.matrix(df[, c("longitude", "latitude")]),
                    miles = FALSE)
nnDmin <- do.call(rbind, 
                  lapply(1:5, function(i) sort(nnDm[, i])[2]))
min(nnDmin)  # e.g., 0.933 km

df_fixed <- st_as_sf(df, 
                     coords = c("longitude", "latitude"), 
                     crs = 4326)

world <- annotation_borders(database = "world", 
               colour = "gray80", 
               fill = "gray80")

simple_map <- ggplot() +
                world + 
                geom_sf(data = df_fixed, 
                        color = "blue") +
                coord_sf(xlim = c(min(df$longitude) - 3, 
                                  max(df$longitude) + 3),
                         ylim = c(min(df$latitude) - 3,
                                  max(df$latitude) + 3)) +
                xlab("Longitude") + 
                ylab("Latitude") +
                annotation_scale() +
                annotation_north_arrow(location = "tl", 
                                       height = unit(1, "cm"),
                                       width = unit(1, "cm"))

write.csv(df, "data/pt_occs_clean/corymbosa_2026_02_03_cleaned.csv", row.names = FALSE)

leaflet(df_fixed) %>%
  addMarkers(label = paste0(df$longitude, ", ", df$latitude)) %>%
  addTiles()
```

## pull occurrences for all taxa

```{julia}
# load in dictionary
synonymdict = load("data/synonymdict.jld2")
# load in dataframe
taxa_traits = DataFrame(CSV.File("data/traits_species_cleaned-2026_02_09.csv"))

# Track failures and no-data taxa
failed_taxa = String[]
successful_taxa = String[]
no_data_taxa = String[]

for taxa in keys(synonymdict)
    synlist = [syn[1] for syn in synonymdict[taxa]]
    filename = replace(taxa, " " => "_")
    filepath = "data/pt_occs_raw/$(filename)-2026_02_18.csv"
    
    # Check if file already exists
    if isfile(filepath)
        println("  Skipping - file already exists")
        push!(successful_taxa, taxa)
        continue
    end
    
    @rput synlist
    @rput filepath
    
    try
        R"gators_download(synonyms.list = synlist, write.file = TRUE, filename = filepath)"
        println("  Success!")
        push!(successful_taxa, taxa)
        sleep(3)
        
    catch e
        error_msg = string(e)
        
        if occursin("No records found", error_msg)
            println("  No occurrence data found for this taxon")
            push!(no_data_taxa, taxa)
            sleep(1)
        else
            println("  ERROR: $e")
            push!(failed_taxa, taxa)
            sleep(10)
        end
    end
end

if !isempty(no_data_taxa)
    println("\nTaxa with no occurrence records:")
    for taxa in no_data_taxa
        println("  - $taxa")
    end
end

if !isempty(failed_taxa)
    println("\nFailed taxa (errors):")
    for taxa in failed_taxa
        println("  - $taxa")
    end
end
```

no failed taxa, no no-data taxa. cool!

batch cleaning for all species:
```{julia}
# Load homonyms file to check for bad authorities
homonyms = DataFrame(CSV.File("data/homonyms.csv"))
false_entries = homonyms[homonyms.accepted .== false, :]
true_entries = homonyms[homonyms.accepted .== true, :]

# Create a dictionary for quick lookup: taxon name -> bad authority
bad_authorities = Dict{String, String}()
for row in eachrow(false_entries)
    bad_authorities[row.scientificName] = row.authority
end

# Create reverse lookup: 
# If validName from FALSE entry is in our dataset, we need to filter OUT the TRUE entry
# BUT only if validName != scientificName (skip self-referential cases)
# Example: "Calceolaria glandulosa Benth." (FALSE) points to "Calceolaria nitida"
#          "Calceolaria nitida" IS in our dataset
#          So when processing "Calceolaria nitida", filter OUT "Calceolaria glandulosa Poepp. ex Benth." (a valid name!)
authorities_to_exclude = Dict{String, Vector{String}}()  # validName -> [authorities to exclude]

dataset_taxa = Set(keys(synonymdict))

for row in eachrow(false_entries)
    valid_name = row.validName
    sci_name = row.scientificName
    
    # Skip if no validName
    if ismissing(valid_name) || valid_name == ""
        continue
    end

    # Skip self-referential cases (validName == scientificName)
    if valid_name == sci_name
        continue
    end
    
    # Check if this validName is in our dataset
    if valid_name in dataset_taxa
        # Find the TRUE entry for this scientificName
        true_match = filter(r -> r.scientificName == sci_name && r.accepted == true, homonyms)
        
        if nrow(true_match) > 0
            true_authority = true_match[1, :authority]
            
            # When processing validName, exclude the TRUE authority
            if !haskey(authorities_to_exclude, valid_name)
                authorities_to_exclude[valid_name] = String[]
            end
            push!(authorities_to_exclude[valid_name], true_authority)
        end
    end
end

println("Taxa with bad authorities to filter: $(length(bad_authorities))")
println("Taxa needing reverse exclusion: $(length(authorities_to_exclude))")
for (valid_name, exclude_auths) in authorities_to_exclude
    for auth in exclude_auths
        println("  $valid_name will EXCLUDE authority: '$auth'")
    end
end

# Initialize tracking dataframe
filtering_stats = DataFrame(
    taxa = String[],
    raw = Int[],
    after_taxa_clean = Int[],
    after_remove_duplicates = Int[],
    removed_bad_authorities = Int[],
    removed_reverse_exclusions = Int[],
    rows_missing_coords = Int[],
    rows_with_coords = Int[],
    after_locality_filt = Int[]
    #after_remove_flagged = Int[],
    #after_1ptper_pixel = Int[]
)

for taxa in keys(synonymdict)    
    # get the synonym list for this species
    synlist = [syn[1] for syn in synonymdict[taxa]]
    filename = replace(taxa, " " => "_")
    filepath = "data/pt_occs_raw/$(filename)-2026_02_18.csv"
    
    # Check if this taxon has a bad authority to filter
    bad_auth = get(bad_authorities, taxa, nothing)
    
    # Check if this taxon needs reverse exclusion (filter out TRUE authorities)
    exclude_auths = get(authorities_to_exclude, taxa, String[])

    # pass to R
    @rput filepath
    @rput synlist
    @rput bad_auth
    @rput exclude_auths
    
    # run R cleaning pipeline with row tracking
    R"""
    df <- read.csv(filepath)
    nrow_raw <- nrow(df)
    
    df <- taxa_clean(df,
                     synonyms.list = synlist, 
                     taxa.filter = "fuzzy",
                     accepted.name = synlist[1])
    nrow_taxa_clean <- nrow(df)
    
    # remove duplicates between GBIF and iDigBio
    df <- remove_duplicates(df, 
                            remove.unparseable = TRUE)
    nrow_remove_dup <- nrow(df)

    # Filter out rows with bad authorities (for homonym cases)
    nrow_removed_bad_auth <- 0
    if (!is.null(bad_auth) && bad_auth != "") {
        df_before <- nrow(df)
        df <- df[!grepl(bad_auth, df$scientificName, fixed = TRUE), ]
        nrow_removed_bad_auth <- df_before - nrow(df)
    }
    
    # Filter out rows with reverse exclusion authorities
    # Example: for Calceolaria nitida, exclude "Poepp. ex Benth." 
    nrow_removed_reverse <- 0
    if (length(exclude_auths) > 0) {
        df_before <- nrow(df)
        for (auth in exclude_auths) {
            df <- df[!grepl(auth, df$scientificName, fixed = TRUE), ]
        }
        nrow_removed_reverse <- df_before - nrow(df)
    }
    
    # Separate rows with NA lat/long for potential georeferencing
    df_na <- df[is.na(df$latitude) | is.na(df$longitude), ]
    df_clean <- df[!is.na(df$latitude) & !is.na(df$longitude), ]
    
    nrow_na <- nrow(df_na)
    nrow_with_coords <- nrow(df_clean)
    
    df_clean <- basic_locality_clean(df_clean, 
                                remove.zero = TRUE, 
                                precision = TRUE, 
                                digits = 2, 
                                remove.skewed = TRUE)
    nrow_locality <- nrow(df_clean)
    
    #df_clean <- process_flagged(df_clean, 
    #                       interactive = FALSE, 
    #                       scientific.name = "accepted_name")
    #nrow_flagged <- nrow(df_clean)
    
    #df_clean <- one_point_per_pixel(df_clean)
    #nrow_pixel <- nrow(df_clean)
    """
    
    # Get row counts from R
    @rget nrow_raw
    @rget nrow_taxa_clean
    @rget nrow_remove_dup
    @rget nrow_removed_bad_auth
    @rget nrow_removed_reverse
    @rget nrow_na
    @rget nrow_with_coords
    @rget nrow_locality
    #@rget nrow_flagged
    #@rget nrow_pixel
    
    # add to tracking dataframe
    push!(filtering_stats, (taxa, nrow_raw, nrow_taxa_clean, nrow_remove_dup, 
                            nrow_removed_bad_auth, nrow_removed_reverse,
                            nrow_na, nrow_with_coords, nrow_locality))
    #push!(filtering_stats, (taxa, nrow_raw, nrow_taxa_clean, nrow_remove_dup, 
    #                         nrow_locality, nrow_flagged, nrow_pixel))
    
    println("  $taxa: Raw=$nrow_raw → Taxa clean=$nrow_taxa_clean → Remove dup=$nrow_remove_dup")
    if nrow_removed_bad_auth > 0
        println("    ⚠️  Removed $nrow_removed_bad_auth rows with bad authority: '$bad_auth'")
    end
    if nrow_removed_reverse > 0
        println("    ⚠️  Removed $nrow_removed_reverse rows with excluded authorities: $(join(exclude_auths, ", "))")
    end
    println("    Split: $nrow_with_coords with coords, $nrow_na missing coords")
    
    # save cleaned df with coordinates
    outfile = "data/pt_occs_clean/$(filename)-2026_02_18_cleaned.csv"
    @rput outfile
    
    R"""
    write.csv(df_clean, 
              outfile, 
              row.names = FALSE)
    """
    
    println("  Saved to: $outfile")
    
    # save rows with NA coordinates (if any exist)
    if nrow_na > 0
        outfile_na = "data/pt_occs_to_georeference/$(filename)-2026_02_18_no_coords.csv"
        @rput outfile_na
        
        R"""
        write.csv(df_na, 
                  outfile_na, 
                  row.names = FALSE)
        """
        println("  Saved $nrow_na rows needing georeferencing to: $outfile_na")
    end
    
    R"rm(df, df_clean, df_na, outfile)"
end

println("\nProcessing complete!")
```

### check minimum number of occurrences
```{julia}
# add calculated columns for seeing filtering before and after nrow
filtering_stats.taxa_clean_removed = filtering_stats.raw .- filtering_stats.after_taxa_clean
filtering_stats.duplicates_removed = filtering_stats.after_taxa_clean .- filtering_stats.after_remove_duplicates
filtering_stats.total_removed = filtering_stats.raw .- filtering_stats.after_remove_duplicates
filtering_stats.percent_retained = round.((filtering_stats.after_remove_duplicates ./ filtering_stats.raw) .* 100, digits=1)

# show taxa with filtering
filtered_taxa = filtering_stats[filtering_stats.total_removed .> 0, :]
println("\n" * "="^60)
println("Taxa with records removed ($(nrow(filtered_taxa)) taxa):")
println(filtered_taxa)

# save the stats
#CSV.write("data/filtering_stats_2026_02_18.csv", filtering_stats)
#println("\nFiltering statistics saved to data/filtering_stats_2026_02_18.csv")
```

let's look through some of these:
1. Calceolaria alba
> this is because many rows were Calceolaria thyrsiflora for some reason? can't see why. so good filtering here.
   
2. Calceolaria collina subsp. conferta
> removing Calceolaria fulva f. viscosa --> good filtering, this is a synonym of Calceolaria fulva.

3. Calceolaria corymbosa subsp. corymbosa
> removes TONS of Calceolaria corymbosa (without any infraspecific label), so good job.

4. Calceolaria uniflora
> removes many Calceolaria virgata (which is a valid taxon), so good job.

before I go down rabbit hole of attempting strict filtering (I might have to use species authorities? but how?),
let's see how many taxa do not have enough occurrences.

```{julia}
toosmall = nrowcomp[ ( nrowcomp.nrow_before .< 25), :]
histogram(nrowcomp.nrow_before)
```

## visualize occurrences before & after filtering

```{julia}
```

fixits:
- edge cases for members in my `homonyms.csv` file --> 
```{julia}
homonymsa = intersect(taxa_traits.scientificName, homonyms.scientificName)
homonymsb = intersect(taxa_traits.scientificName, homonyms.validName)
homonyms_indataset = union(homonymsa, homonymsb)
```

there's 13 of these, we'll need to be extra careful and double check:
  1. occurrences of the synonym/homonym without species authority
  2. look at overlap of range between synonyms

- create a dictionary of native ranges? don't think gatoRs does anything with native range --> just relies on cultivated tag?
  - pull native range dictionary from POWO --> create dictionary
  - create function to pull botanical country codes from shapefile (?)
  - use `CoordinateCleaner` to pull weird occurrences, and use native range shapefiles (?)
- get summary stats for cultivated vs. inat vs. etc
